Collaborative Visual Analysis on the Web with RCloud

### 2015-03-31 discussion.tex
more tweaking

{\bf Shared artifacts and artifact histories} are a central feature of
RCloud, through its shared workbooks of experiments and analyses.  We
observed that hackers readily shared work through RCloud. They used it to
discuss ideas and results in data analysis, to ask for help with
programming problems, and to publish results by sharing URLs.  Artifact
histories can be accessed through the notebook tree or through GitHub's web
interface.

{\bf Modularity and granularity.} The ability to partition work into
independent units (modularity) is a key to working productively in teams.
It is good if the units can be kept small, so team members can realize
benefits at least proportional to the work on the units (granularity).
RCloud's notebook and versioning capabilities allow work to be divided
into units as fine as the underlying language allows, and versioning
encourages making incremental changes at low cost and without disrupting
the work of others. \gordon{But this design choice means that developing
larger, multi-page apps is less well supported.}

@@

{\bf Discussion.} Annotation and commenting was another central goal.
Commenting is supported through github, but our hackers found it
awkward and did not exercise it as much as we expected.
An interesting question is, to what extent should application users
be able to make annotations in published notebooks without coding
and being exposed to the hackers's view? The design of
more elaborate, integrated annotation remains as future work.

{\bf Content export} is not a capability we aimed at supporting,
and R already has many packages for this. Instead, our goal was to
encourage ways of working without needing as much explicit export.
We did assume that RCloud notebooks should play well in the ecosystem
of other tools, and this worked as intended, though we learned it also
creates a temptation to absorb features almost without limit. (Should
RCloud provide templates for personal web pages and graphical
editing? Should it incorporate an email client and emacs?
Should it push results to Facebook?) Experience shows that interfaces
to external systems and other features almost always win over architectural
purity in the real world. \gordon{If you don't export, doesn't that mean
you have to support every possible operation downstream? Due to popular demand,
we did add plot export.}

{\bf Social-psychological incentives} and {\bf voting and ranking} are supported
through starring. This mechanism was employed often by our test
audience. \gordon{actually, the notebook tree quickly became so unwieldy that
no one uses this.} An obvious next step would be to enhance recommendations using
relationships discovered by static and dynamic code analysis. This may be
considered both within and across collections of scripts (the latter being
similar to VisTrails' enhanced recommendations by clustering multiple
workflows).  It seems valuable to know which packages are frequently used
together, or appear in proximity to a certain record types or data feeds.  In
general, trivial or passive mechanisms to collect data for recommendations are
essential.

@@

{\bf Versioning.}
Some important aspects of the environment are more difficult to manage
in RCloud than in conventional systems. RCloud does not explicitly
separate the development and deployment environments. More than that,
ever version of every prototype is shared by default. Although this encourages
collaboration on work in progress it also quickly exposes errors such as
misconfiguration, mismatches between versions of packages, and programming
errors that can unintentionally affect production websites. Such errors
are difficult to completely avoid, but the power of convenient sharing is often
worth it (as proven true with distributed systems in general).
Still, effective control over versioning when sharing problem needs more
attention if RCloud would be scaled up to a wider number of users lacking
informal channels to coordinate work.

@@

{\bf Exploration is hard.}
R is a popular language for exploratory data analysis, and we built RCloud
as a way for data scientists to seamlessly transition from EDA to sharing
and presentation. But the combined notebook and command line interface we designed interferes with the exploration aspect.

>>>>

{\bf Shared artifacts and artifact histories} are a central feature of
RCloud, through its shared workbooks of experiments and analyses.  We
observed that hackers readily share work through RCloud. They use it to
demonstrate techniques, share results with peers and managers, and transfer algorithms
to other groups.  Artifact histories can be accessed through the notebook
tree or through GitHub's web interface.

{\bf Modularity and granularity.} The ability to partition work into
independent units (modularity) is a key to working productively in teams.
It is good if the units can be kept small, so team members can realize
benefits at least proportional to the work on the units (granularity).
RCloud's notebook and versioning capabilities allow work to be divided
into units as fine as the underlying language allows, and versioning
encourages making incremental changes at low cost and without disrupting
the work of others.

@@

{\bf Discussion.} Annotation and commenting was another central goal.
Commenting is supported through GitHub, but our hackers found it
awkward and did not exercise it as much as we expected.
An interesting question is, to what extent should application users
be able to make annotations in published notebooks without coding
and being exposed to the hackers's view? The design of
more elaborate, integrated annotation remains as future work.

{\bf Content export} is not a capability we aimed at supporting,
and R already has many packages for this. Recently, due to popular
demand, we added user interface support for exporting plot images.
We assume that RCloud notebooks should play well in the ecosystem
of other tools, but sometimes it is difficult to know whether adding
a compatibility feature will offload complexity or bring more in.

{\bf Social-psychological incentives} and {\bf voting and ranking} are
supported through starring and forking. These mechanisms are employed often
on the platform. An obvious next step would be to enhance recommendations
using relationships discovered by static and dynamic code analysis. This
may be considered both within and across collections of scripts (the latter
being similar to VisTrails' enhanced recommendations by clustering multiple
workflows).  It seems valuable to know which packages are frequently used
together, or appear in proximity to a certain record types or data feeds.
In general, trivial or passive mechanisms to collect data for
recommendations are essential.

@@

{\bf Versioning.}
Some important aspects of the environment are more difficult to manage
in RCloud than in conventional systems. RCloud does not explicitly
separate the development and deployment environments. More than that,
every version of every prototype is shared by default. Although this encourages
collaboration on work in progress it also quickly exposes errors such as
misconfiguration, mismatches between versions of packages, and programming
errors that can unintentionally affect production websites. Such errors
are difficult to completely avoid, but the power of convenient sharing is often
worth it (as proven true with distributed systems in general).
Still, effective control over versioning when sharing problem needs more
attention if RCloud would be scaled up to a wider number of users lacking
informal channels to coordinate work.

@@

{\bf Exploration is hard.}
R is a popular language for exploratory data analysis, and we built RCloud
as a way for data scientists to seamlessly transition from EDA to sharing
and presentation. But the combined notebook and command line interface we
designed interferes with the use of R for exploration.
